MLH1 Results Outline

### 2020-02-26 doc/Results.rmd
working from bottom up

DO THE HIGH QUALITY dataset tests

@@

@@

@@

#SKIVE
#Q1_tot.sc_t.test.SKIVE <- t.test((totSC.mouse.av.table %>% filter(sex == "male") #%>% filter(strain == "SKIVE") )$mean.skel,
#           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == #"SKIVE") )$mean.skel, paired = FALSE, alternative = "greater")

#error not enought female observations

@@

The most averages for total SC-AE area (skeletonizes) are significantly longer in females for all strains (p=
`r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, and `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`
for the _domesticues^WSB^ _, _domesticues^G^ _, and _domesticues^LEW^ _ respectively).

Some of the p values for PWD and MSM are much closer (less significant), ( p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, and `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )` for the _musculus^PWD^ _, _musculus^KAZ^ _, and _molossinus^MSM^_ respectively)

@@

The mouse means for total SC in domesticus strains are significant shorter compared to Musc (t.test; p= `r Q1.male.Dom.Musc.t.test$p.value`) and Molossinus (t.test; p = `r Q1.male.Dom.Mol.t.test$p.value`, indecating evolution (divergence) of SC length and (chromatin compaction) <this will be expanded on n later sections>
-- we follow up with models (which focus on sex differences).

<explain why applying models.  We then use models to examine/ think about the evolution>

The above t.tests were done to establish the sex differences for SC-AE (chromatin condensation), we now apply models (for understanding within an evolutionary framework) -- we apply the same models to both mean mouse total sc and mouse mean short bivalent.

@@

**M2.Linear Model** $$mouse \ average \ SC \ metric ~=~ sex * strain + \varepsilon $$

@@

For the mixed model run on the total sc metric, there are more significant effects for strain and subsp for the mixed model of mouse.av total sc.

The three fixed effects and one random effect were significant (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, 
sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`).

@@

Mixed models for the sex differences in mouse averages for short bivalents Sex is the most significant factor for the mixed model of mean short bivalents (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`). The interaction (subspecies by sex) effect also significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`).

These effects were followed up the glms, again applied to both mouse means for short biv SC lengths and total sc area per cell.

@@

The following effects had significant p values for the **total SC** from the above model

subspecies musc p = `r sum.Q1.glm_totSC_M2$coefficients[,4][2]`, subspecies mol p =`r sum.Q1.glm_totSC_M2$coefficients[,4][3]`, sex male  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][8]`, strain lew p =`r sum.Q1.glm_totSC_M2$coefficients[,4][5]`, strain g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][4]`,
<this is the smallest p vale, elevated female total, driven by the high means in females>

(subsp*sex) mol.male  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][10]`, sex.g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][11]`, lew.male p =`r sum.Q1.glm_totSC_M2$coefficients[,4][12]`

<!--
for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _
-->

Fr the **short bivalent** sc the model results are listed below. From this (M2) model sex effect, male is most significant (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]`).

while the skive strain effect (glm; p=`r summ.Q1.glm_short.SC_M2$coefficients[,4][7]`) and and SKIVE*male (glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][16]`) are slightly significant.

In the M3 model (strains and sex as fixed effects), the Sex effect (male) is the only significant p value from this model (glm; p= `r summ.Q1.glm_short.SC_M3$coefficients[,4][14]`).

@@

@@

# Q2 Normalized Single CO positions

<!-- motivation / background of the sex differences, males more terminall in CO placement compared to females, now we ask if there are significant differences in the placement of single foci along bivalents  -->
<!-- set up chunks here -->
<!-- plots /figures below-->

@@

<not sure what the beenfit is to test for chrm effectss>

@@

The previous section found that the single crossover landscape is significantly different between males and females.
In this section we focus on differences between the high and the low rec male strains
In order to test if if there are distinguishing meterics for these two groups.

@@

**The figure above shows, WSB has significantly more terminal positioning, while MOLF has significantly more medial position.**

T tests of the single crossover positions reveal that the normalized single foci positions are not significantly different between the total pooled high and low rec male strains (t.test; p = `r Rec.group_BivData_F1post_Ttest$p.value`).

<further tests -- within the subspecies, musculus and molossinus?

The high rec musc strains, _musculus^PWD^_ and _musculus^SKVIE^_, do not have signifcantly different single foci landscapes from the other musc low rec strains (ttest; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value`).

While _molossinus^MOLF^_ is significantly more centrally positioned compared to _molossinus^MSM^_ (t.test; p = `r Q2.1COpos_mus.av_Mol_ttest$p.value`). This result is somewhat puzzeling and it doesn't match up with the strains with high recombination rates.

 <BUT this might be due to sampling for MSM -- but it doesn't seem to be the case in MOLF, 
 (MSM and MOLF have different proportions of chromosome classes -- (MSM has the smallest propotion of 1COs, across the strains, BUT the mean and distributions for MSM of single foci positions are equivilent across all the strains.>

@@

in addition models which use mouse average of foci position as a dependat variable, we decided to run models for the single bivalent level, and which include chromosome length as a fixed effect.

in the above chunk I tried making another glm's which uses the single bivalent measures and include chromosome size for the switching the model to include interaction (instead of additive), all of the pvalues get lower. .. the chromosome length has the most significant effect, LEW by chrm effect are significant

In the reduced model, chrm length is the most significant, strain SKIVE, MOLF, G and LEW are significant strain effects

**For the M2 (subsp)**
There are differences for the dom strains: G strain(p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][4]`).

LEW (slight) (p =`r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][5]`).

But big effects from Mol (p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][3]`).

and MSM strain (p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][7]`)

**For M3 (all strains)**

@@

strain KAZ (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][8]`).

from the main figure above, you can see that G and MOLF 1CO foci positions are the most different

<no prediction for the strains within subspecies>

For MOLF 326 1CO bivalents,  79, 126, 62, 59 1CO observations across the 4 MOLF mice

there are only 77 MSM 1COs across 4 mice in MSM (MSM is less central than MOLF)
(motivation for MOLF female observations and more male MSM biv measures)

@@

<I ran another model that used the single bivalents observations to predict the normalize foci position with fixed effects of chromosome length, subspecies, strian and the interaction effects.

**M.X. Linear Model** $$1CO \ Normalized \ Position ~=~ chromosomeLength * subsp * strain + \varepsilon $$>

@@

<for the first model,

chromosome length `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][2]`

@@

-chrm length LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][12]`>

males from WSB and MOLF strains have 'strang'unique patterns for the single foci positions, which don't fit with predicted patterns from the genome wide recombination rate variation

There are significant difference across the 1CO positions, yet they don't match fit any of the predictions. there are also complications due to the chromsome class proportions that

@@

<glms>

The following model framework was applied to further investigate these patterns using mouse averages of IFD^raw^ and IFD^PER^.

@@

<when the M2 model was run within subspecies, the results generally confirmed our predictions; no sig values for Dom, PWD and SKIVE have longer IFDs than other musc strains and MSM is longer than MOLF>

@@

@@

<!-- unused chunkcs -->

@@

<!-- Whole Cell Manual Comparison -->

@@

# Note on Heterochiasmy Definition

@@

>>>>

@@

#ALL
Q1_tot.sc_t.test.all <- t.test((totSC.mouse.av.table %>% filter(sex == "male") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female"))$mean.skel)

#dom
Q1_tot.sc_t.test.DOM <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(subsp == "Dom") )$mean.skel)

@@

#Musc
Q1_tot.sc_t.test.MUSC <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )$mean.skel,
                           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(subsp == "Musc") )$mean.skel)

@@

#SKIVE
#Q1_tot.sc_t.test.SKIVE <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$mean.skel,
#           (totSC.mouse.av.table %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$mean.skel, paired = FALSE, alternative = #"greater")
#cell level total SC
#MLH1.merge.Skel   $skel_size
#error not enought female observations, ONLY 1 female SKIVE mouse

Q1_tot.scBIV_t.test.SKIVE <- t.test((MLH1.merge.Skel %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$skel_size,
           (MLH1.merge.Skel %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$skel_size   )

@@

In all of the strains, the female mouse mean total SC per cell was significantly greater in females than males (ttest; WSB p = `r format(Q1_tot.sc_t.test.WSB$p.value, scientific = F )`, G p = `r format(Q1_tot.sc_t.test.G$p.value, scientific= F )`, LEW p = `r format(Q1_tot.sc_t.test.LEW$p.value, scientific= F )`,  PWD p = `r format(Q1_tot.sc_t.test.PWD$p.value,scientific = F )`, KAZ p = `r format(Q1_tot.sc_t.test.KAZ$p.value, scientific = F )`, MSM `r format(Q1_tot.sc_t.test.MSM$p.value, scientific = F )`

for the _domesticues^WSB^ _, _domesticues^G^ _, and _domesticues^LEW^ _ respectively), for the _musculus^PWD^ _, _musculus^KAZ^ _, and _molossinus^MSM^_ respectively)

@@

<!-- Dom males are shorter than this set of high rec mice
The mouse means for total SC in domesticus strains are significant shorter compared to Musc (t.test; p= `r Q1.male.Dom.Musc.t.test$p.value`) and Molossinus (t.test; p = `r Q1.male.Dom.Mol.t.test$p.value`, indecating evolution (divergence) of SC length and (chromatin compaction) <this will be expanded on n later sections>-- we follow up with models (which focus on sex differences).  -->

<**t.test results for total.SC**
all values (F vs M) (p= `r Q1_tot.sc_t.test.all$p.value`)
Dom (F vs M) -- (ttest; p=`r Q1_tot.sc_t.test.DOM$p.value`)
Musc (F vs M) -- (ttest; p=`r Q1_tot.sc_t.test.MUSC$p.value`)  (note, missing SKIVE, not enough female measures)  
MSM (F vs M) -- (ttest; p= `r Q1_tot.sc_t.test.MSM$p.value`)>

<explain why applying models.  We then use models to examine/ think about the evolution>
The above t.tests were done to establish the sex differences for SC-AE (chromatin condensation), we now apply models (for understanding within an evolutionary framework) -- we apply the same models to both mouse mean short bivalent length and mean mouse total SC.

@@

@@

@@

Mixed models for the sex differences in mouse in the mixed models using mean **short bivalents** length as the dependant variable, sex is the most significant effect (LTR; p =`r Q1lmer.MM_SC_short_results$sex.results`). The interaction (subspecies by sex) effect was also slightly significant (p = `r Q1lmer.MM_SC_short_results$interaction.pval`).

Subsp and sex had the most significant p values, --- For the mixed model run on the total sc metric, there are more significant effects for strain and subsp for the mixed modelusing mean **total sc** per cell as the dependant variable, the three fixed effects and one random effect were significant (LTR; subsp p= `r Q1.lmer.totSC_results$sub.sp`, sex p = `r Q1.lmer.totSC_results$sex.results`, and interaction p = `r Q1.lmer.totSC_results$interaction.pval`, (random LRT; strain p = `r Q1.lmer.totSC_results$rand.pval`).

(these results from the mixed models -- are similar to the mean MLH1 per cell results -- so we follow up -- with tests of significant strain effects).

@@

In linear models (M2 and M3) for **mean short bivalent length**, male is most significant (M2 glm; p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][8]` and M3 glm; p = `r summ.Q1.glm_short.SC_M3$coefficients[,4][8]`). In the linear model M2 the skive strain and strain by sex effects are slightly significant (glm; p=`r summ.Q1.glm_short.SC_M2$coefficients[,4][7]`, p =`r summ.Q1.glm_short.SC_M2$coefficients[,4][14]` respectively), however these could be due to low sampling of skive females.

The following effects had significant p values for the **total SC** from the above model (for the linear models M2 and M3), 
subspecies _musculus_ (glm M2; p = `r sum.Q1.glm_totSC_M2$coefficients[,4][2]`), subspecies _molossinus_ (glm M2; p =`r sum.Q1.glm_totSC_M2$coefficients[,4][3]`), sex male (glm M2; p =`r sum.Q1.glm_totSC_M2$coefficients[,4][8]`), strain lew p =`r sum.Q1.glm_totSC_M2$coefficients[,4][5]`, strain g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][4]`,

For M3 -- many strain effects are significant, but the largest are sex (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][8]`)
but also some strain effects, G strain (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][2]`), LEW strain (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][3]`), PWD strain (glm M3; p = `r sum.Q1.glm_totSC_M3$coefficients[,4][4]`).

Supporting the strong sex difference and also indicating that SC length has evolved across strains within this panel. when chromatin condensation is summarized / quantified at this level. (can't exclude the possibility that the XY morphology is different between these strains)

<!--
(subsp*sex) mol.male  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][10]`, sex.g  p =`r sum.Q1.glm_totSC_M2$coefficients[,4][11]`, lew.male p =`r sum.Q1.glm_totSC_M2$coefficients[,4][12]`
for the _mmdomesticues^WSB^ _, _mmdomesticues^G^ _, and _mmdomesticues^LEW^ _
-->

@@

```{r Q1.SC.Q12, echo=FALSE}

#make comparisons for the high quality cells

#get Q12 cells from MLH1.merge.Skel
#calq the mouse averages

MLH1.merge.Skel$quality <- as.numeric(MLH1.merge.Skel$quality)

totSC_Q12_mouse.av.table <- ddply(.data=MLH1.merge.Skel %>% filter(quality < 3), 
                 .(mouse),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),

totSC_Q12_mouse.av.table <- add_strain(totSC_Q12_mouse.av.table)
totSC_Q12_mouse.av.table <- add_sex(totSC_Q12_mouse.av.table)
totSC_Q12_mouse.av.table <- add_category(totSC_Q12_mouse.av.table)
totSC_Q12_mouse.av.table <- add_subsp(totSC_Q12_mouse.av.table)
totSC_Q12_mouse.av.table <- add_species(totSC_Q12_mouse.av.table)

#unorder the factors
totSC_Q12_mouse.av.table$subsp <- factor(totSC_Q12_mouse.av.table$subsp, ordered = FALSE)
totSC_Q12_mouse.av.table$sex <- factor(totSC_Q12_mouse.av.table$sex, ordered = FALSE)
totSC_Q12_mouse.av.table$strain <- factor(totSC_Q12_mouse.av.table$strain, ordered = FALSE)
totSC_Q12_mouse.av.table$species <- factor(totSC_Q12_mouse.av.table$species, ordered = FALSE)

#add rec group
totSC_Q12_mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC_Q12_mouse.av.table$category), 1, 
                              ifelse(grepl("MSM male", totSC_Q12_mouse.av.table$category), 1,
                                ifelse(grepl("SKIVE male", totSC_Q12_mouse.av.table$category), 1, 0)))

#REMOVE JUVI mice
tot.SC.Q12_juvi.male <- totSC_Q12_mouse.av.table[totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]
totSC_Q12_mouse.av.table <- totSC_Q12_mouse.av.table[! totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]

totSC_Q12_strain.av.table <- ddply(.data=MLH1.merge.Skel %>% filter(quality < 3), 
                 .(strain, sex),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),

)
#totSC_Q12_strain.av.table <- add_strain(totSC_Q12_strain.av.table)
#totSC_Q12_strain.av.table <- add_sex(totSC_Q12_strain.av.table)
#totSC_Q12_strain.av.table <- add_category(totSC_Q12_strain.av.table)
totSC_Q12_strain.av.table <- add_subsp(totSC_Q12_strain.av.table)
totSC_Q12_strain.av.table <- add_species(totSC_Q12_strain.av.table)

#unorder the factors
totSC_Q12_strain.av.table$subsp <- factor(totSC_Q12_strain.av.table$subsp, ordered = FALSE)
totSC_Q12_strain.av.table$sex <- factor(totSC_Q12_strain.av.table$sex, ordered = FALSE)
totSC_Q12_strain.av.table$strain <- factor(totSC_Q12_strain.av.table$strain, ordered = FALSE)
totSC_Q12_strain.av.table$species <- factor(totSC_Q12_strain.av.table$species, ordered = FALSE)

```{r Q1.SC_showQ12Tot.SC, echo=FALSE, warning=FALSE}

Q12.Tot.SC_mouse.av_Dom <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Dom"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(~sex, scales="free", ncol = 3)+

  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Dom"), aes(yintercept=mean.skel) )+

  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,2600 ))+ggtitle("Mouse av. Q12 total SC area")

#Q12.Tot.SC_mouse.av_Dom
#Tot.SC_mouse.av_Dom

Q12.Tot.SC_mouse.av_Musc <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Musc"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(~sex, scales="free", ncol = 3)+

  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Musc"), aes(yintercept=mean.skel) )+

  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")

#Q12.Tot.SC_mouse.av_Musc
#Tot.SC_mouse.av_Musc

Q12.Tot.SC_mouse.av_Mol <- ggplot(data= totSC_Q12_mouse.av.table %>% filter(subsp == "Mol"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(~sex, scales="free", ncol = 3)+

  geom_hline(data =totSC_Q12_strain.av.table %>% filter(subsp == "Mol"), aes(yintercept=mean.skel) )+

  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())+  ylim(c(1000,3000 ))+ggtitle("Mouse av. Q12 total SC area")

Q12.Tot.SC_mouse.av_Mol
Tot.SC_mouse.av_Mol

<Q12 total SC mouse means!! In Dom when running the Q12 SC lengths, the sex difference seems to get larger (male mean total SC gets lower).  In Musc the female mean total sc gets lower (maybe fewer cells (SKIVE is removed)). For Mol / MSM female mean stays about the same, MSM and MOLF means converge around 1600.>

@@

# Q2.1CO rec landscape evolution is decoupled from gwRR evolution

<!-- motivation / background of the sex differences, males more terminall in CO placement compared to females, now we ask if there are significant differences in the placement of single foci along bivalents

Main results:  
1. WSB has the most terminal 1CO rec landscape and MOLF has the most central.
2. Evidence for evolution of the 1CO rec landscape, but it is decoupled from the evolution in gwRR we observe in figure 1
2.5  The normalized 1CO position is not a good metric for distinguishing strains with rapid gwRR evolution

-->
<!-- set up chunks here -->
<!-- plots /figures below-->

@@

@@

@@

The previous section found that the single crossover landscape is significantly different between males and females. In this section we focus on differences between the high and the low rec male strains
In order to test if if there are distinguishing meterics for these two groups. **The figure above shows, WSB has significantly more terminal positioning, while MOLF has significantly more medial position.**

T tests of the single crossover positions reveal that the normalized single foci positions are not significantly different between the total pooled high and low rec male strains (t.test; p = `r Rec.group_BivData_F1post_Ttest$p.value`). The high rec musc strains, _musculus^PWD^_ and _musculus^SKVIE^_, do not have signifcantly different single foci landscapes from the other musc low rec strains (ttest; p = `r Q2.1COpos_mus.av_Musc_ttest$p.value`). While _molossinus^MOLF^_ is significantly more centrally positioned compared to _molossinus^MSM^_ (t.test; p = `r Q2.1COpos_mus.av_Mol_ttest$p.value`). This result is somewhat puzzeling and it doesn't match up with the strains with high recombination rates.

@@

<!--
In the above chunk I tried making another glm's which uses the single bivalent measures and include chromosome size for the switching the model to include interaction (instead of additive), all of the pvalues get lower. .. the chromosome length has the most significant effect, LEW by chrm effect are significant

In the reduced model, chrm length is the most significant, strain SKIVE, MOLF, G and LEW are significant strain effects -->

<!-- Just hold these for reference
**For the M2 (subsp)**
There are differences for the dom strains: G strain(p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][4]`).

LEW (slight) (p =`r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][5]`).  
But big effects from Mol (p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][3]`).  
and MSM strain (p = `r summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4][7]`)  
**For M3 (all strains)**

@@

strain KAZ (p = `r sum.Q2.glm_mus.av_1COpos_M2[,4][8]`).
-->

@@

given that the t.test don't return significant pvalues so we are not motivated to follow up with linear models. Males from WSB and MOLF strains have 'strang'unique patterns for the single foci positions, which don't fit with predicted patterns from the genome wide recombination rate variation

<I ran another model that used the single bivalents observations to predict the normalize foci position with fixed effects of chromosome length, subspecies, strian and the interaction effects.>

<!--
**M.X. Linear Model** $$1CO \ Normalized \ Position ~=~ chromosomeLength * subsp * strain + \varepsilon $$ -->

@@

<!-- hide these linear model results for now
for the first model, 
chromosome length `r summary(chrm.1CO.pos.chrmLength)$coefficients[,4][2]`

@@

-chrm length LEW, `r summary(chrm.1CO.pos.chrmLength_strain)$coefficients[,4][12]` -->

@@

<glms> The following model framework was applied to further investigate these patterns using mouse averages of IFD^raw^ and IFD^PER^.

@@

<!--when the M2 model was run within subspecies, the results generally confirmed our predictions; no sig values for Dom, PWD and SKIVE have longer IFDs than other musc strains and MSM is longer than MOLF-->

@@

<!-- final summary, BE brief-->

BE brief,

final summary, highlight the themes of 1. more variance in female traits (MLH1 foci per cell, rec landscape / interference)

while the male traits -- more focused (less variance in MLH1 counts per cell, shorter chromatin comparction?  bias position, stronger interference)

(should I measure the var of all traits in Q1)

2. rapid evolution of genome wide rec rates can fit into this -

- evolution of more COs requires more DSBs (evidence points to a earlier step which is likely the chromatin compaction (amount of axis))

- the strongest single bivalent predictor/correlate of rapid gwRR for males is interferece strength in 2COs

AND somewhat puzzeling result of (having stronger interference), meaning there is a greater proportion of a bivalent that seperates two foci  (2 foci push to the ends of chrms (maybe decreased centromere supression))

# Blank1

<!-- unused chunks -->

@@

@@

<!-- Whole Cell Manual Comparison

@@

# Blank2
<!-- Note on Heterochiasmy Definition

@@

  -->
